---
title: "Amplicon_analysis"
author: "Luyao Kang"
date: "`r Sys.Date()`"
output:
  html_notebook:
    toc: yes
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = T, echo = T, comment = "#>", message = F, warning = F,
	fig.align = "center", fig.width= 4.5, fig.height = 3, dpi = 150)
```
## Data input 
Set work directory
```{r}
setwd('e:/permafrost/')
wd_16s <- file.path(getwd(),"data/16S")
# if (!dir.exists(wd_16s)) {
#   dir.create(wd_16s)
# }
wd_fun <- file.path(getwd(),"data/metagenome")
save.dir <- file.path(getwd(),"result")
```
Loading packages
```{r}
library(phyloseq)
library(ape)
library(Biostrings)
library(microbiome)
library(tidyverse)
```
Data input
```{r}
source("script/read_data.R")
```

## Alpha diversity
Determine the alpha diversity including **Observed**, **Chao1**, **Shannon** and **Simpson**.
```{r}
alpha_div <- estimate_richness(phylo_rare, measures = c("Observed", "Chao1", 'Shannon', 'Simpson'))
library(picante)
pd <- pd(t(otu), tree, include.root = F) # estimate the phylogenetic diversity
alpha_div <- cbind(Layers = metadata$Layer, alpha_div, Faith = pd$PD,
                   Evenness = alpha_div$Shannon/log(alpha_div$Observed)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
```
Box plot for evenness index using ggplot2
```{r}
library(ggpubr)
library(ggplot2)
my_comparisons <- list( c("SUR", "SUB"), c("SUB", "PL"), c("SUR", "PL"))
p <- ggplot(alpha_div, aes(x = Layers, y = Evenness)) + 
  geom_boxplot(width = 0.5, aes(fill = Layers)) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_compare_means(comparisons = my_comparisons, paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  #scale_fill_manual(values= cols)+
  labs(x = 'Layers', y = 'Evenness', fill='Layers') +
  theme_bw() +
  theme(axis.title = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank())
p
```
Box plot for faith index using ggplot2
```{r}
p <- ggplot(alpha_div, aes(x = Layers, y = Faith)) + 
  geom_boxplot(width = 0.5, aes(fill = Layers)) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_compare_means(comparisons = my_comparisons,  paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  #scale_fill_manual(values= cols)+
  labs(x = 'Layers', y = 'Faith index', fill='Layers') +
  theme_bw() +
  theme(axis.title = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank())
p
```
## Beta diversity
```{r}
library(vegan)
library(pairwiseAdonis)
## taxonomy
### PERMANOVA test
tax_dist <- as.matrix(vegdist(t(otu), "bray" ))
adonis2(tax_dist ~ Layer, data = metadata)
```
Pairwise adonis test
```{r}
set.seed(123)
pairwise.adonis(tax_dist, metadata$Layer)
```
### PCoA plot with bray-curties as distance
```{r}
ord.tax <-  cmdscale(tax_dist,  k = 2, eig = T, add = T)
pcoa_tax_plot <- data.frame(Layers = metadata$Layer, scores(ord.tax)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Dim1, y = Dim2, shape = Layers, color = Layers)) + 
  geom_point(size = 1, alpha = 0.8) + 
  stat_ellipse(geom = "polygon", aes(fill = Layers), alpha = 0.2, show.legend = FALSE, level = 0.95) +
  scale_colour_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  labs(x=paste("PCoA1 (", format(100 * ord.tax$eig[1] / sum(ord.tax$eig), digits = 3), "%)", sep = ""),
       y=paste("PCoA2 (", format(100 * ord.tax$eig[2] / sum(ord.tax$eig), digits = 3), "%)", sep = "")) +
  theme(axis.title = element_text(size = 8, colour = "black"),
        axis.text = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        panel.grid = element_blank(),
        panel.background = element_blank(), 
        panel.border = element_rect(fill = NA, colour = "black"))
# ggsave(file.path(save.dir, "./figs/beta/PCoA_tax_bray.pdf"),
#        pcoa_tax_plot, width = 89, height = 59, units = "mm")
pcoa_tax_plot
```
### Difference in taxonomic variance among layers
```{r}
beta_tax_plot <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(tax_dist, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Layers, y = distance)) + 
  geom_boxplot(width = 0.5, aes(fill = Layers)) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_compare_means(comparisons = my_comparisons,  paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  #scale_fill_manual(values= cols)+
  labs(x = 'Layers', y = 'Taxonomic variance', fill='Layers') +
  theme_bw() +
  theme(axis.title = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank())
# ggsave(file.path(save.dir, "./figs/beta/tax_variance.pdf"),
#        beta_tax_plot, width = 89, height = 89, units = "mm")
beta_tax_plot
```
## Phylogeny
PERMANOVA test
```{r}
phy_dist <- as.matrix(UniFrac(phylo, weighted = TRUE, normalized = TRUE, parallel = T, fast = TRUE))
adonis2(phy_dist ~ Layer, data = metadata)
```
Pairwise adonis test
```{r}
set.seed(123)
pairwise.adonis(phy_dist, metadata$Layer)
```
### PCoA plot with unweighted UniFrac as distance
```{r}
ord.phy <- ordinate(phylo, method = "PCoA", distance = "unifrac", weighted = TRUE)
PCoA_phy_plot <- plot_ordination(phylo, ord.phy, color = "layer") + theme(aspect.ratio = 1)
PCoA_unifrac_plot <- data.frame(Layers = metadata$Layer, ord.phy$vectors[, 1:2]) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Axis.1, y = Axis.2, shape = Layers, color = Layers)) + 
  geom_point(size = 1, alpha = 0.8) + 
  stat_ellipse(geom = "polygon", aes(fill = Layers), alpha = 0.2, show.legend = FALSE, level = 0.95) +
  scale_colour_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  labs(x=paste("PCoA1 (", format(100 * ord.phy$values[1, 2], digits = 3), "%)", sep = ""),
       y=paste("PCoA2 (", format(100 * ord.phy$values[2, 2], digits = 3), "%)", sep = "")) +
  theme(axis.title = element_text(size = 8, colour = "black"),
        axis.text = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        panel.grid = element_blank(),
        panel.background = element_blank(), 
        panel.border = element_rect(fill = NA, colour = "black"))
# ggsave(file.path(save.dir, "./figs/beta/PCoA_unifrac_plot.pdf"),
#        PCoA_unifrac_plot, width = 89, height = 59, units = "mm")
PCoA_unifrac_plot 
```

### Determine the betaMNTD
#### Method 1
```{r}
require(picante)
beta.mntd.weighted <- as.matrix(comdistnt(t(data.frame(otu)), cophenetic(tree), abundance.weighted = T))
```
#### Method 2 (time efficiency)
```{r}
library(doParallel)  
cores <- detectCores() - 2  #using multiple cores

p.dist.mat <- cophenetic(tree)
get.presents <- function(x) {
  names(x[x > 0])
}
list.of.names <- apply(otu, 2, get.presents)
Dnn.apply.funtion <- function(x) {
  tmp.function <- function(z) {
    mean(c(apply(p.dist.mat[x, z], MARGIN = 1, min, na.rm = T), 
         apply(p.dist.mat[x, z], MARGIN = 2, min, na.rm = T)))
  }
  lapply(list.of.names, FUN = tmp.function)
}
beta.mntd.weighted <- do.call(cbind, lapply(list.of.names, Dnn.apply.funtion))
```
### Reading beta.mntd.weighted file
```{r}
beta.mntd.weighted <- read.table(file.path(save.dir, './tables/beta.mntd.weighted.txt'),
                                 header = T, row.names = 1)
```
### Difference in MNTD among layers
```{r}
beta_MNTD_plot <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(beta.mntd.weighted, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Layers, y = distance)) + 
  geom_boxplot(width = 0.5, aes(fill = Layers)) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_compare_means(comparisons = my_comparisons,  paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  #scale_fill_manual(values= cols)+
  labs(x = 'Layers', y = 'beta-MNTD', fill='Layers') +
  theme_bw() +
  theme(axis.title = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank())
# ggsave(file.path(save.dir, "./figs/beta/beta_mntd.pdf"),
#        beta_MNTD_plot, width = 89, height = 89, units = "mm")
beta_MNTD_plot
```
## Functional analysis
### PERMANOVA test
```{r}
fun_dist <- as.matrix(vegdist(t(ko_tpm_table), "bray" ))
adonis2(fun_dist ~ Layer, data = metadata)
```
```{r}
set.seed(123)
pairwise.adonis(fun_dist, metadata$Layer)
```
### PCoA plot
```{r}
ord.fun <-  cmdscale(fun_dist,  k = 2, eig = T, add = T)
library(ggplot2)
pcoa_fun_plot <- data.frame(Layers = metadata$Layer, scores(ord.fun)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Dim1, y = Dim2, shape = Layers, color = Layers)) + 
  geom_point(size = 1.5, alpha = 0.8) + 
  labs(x=paste("PCoA1 (", format(100 * ord.fun$eig[1] / sum(ord.fun$eig), digits = 3), "%)", sep = ""),
       y=paste("PCoA2 (", format(100 * ord.fun$eig[2] / sum(ord.fun$eig), digits = 3), "%)", sep = "")) +
  theme(axis.title = element_text(size = 6, colour = "black"),
        axis.text = element_text(size = 5, colour = "black"),
        legend.title = element_text(size = 6),
        legend.text = element_text(size = 5),
        legend.key = element_blank(),
        panel.grid = element_blank(),
        panel.background = element_blank(), 
        panel.border = element_rect(fill = NA, colour = "black"))
# ggsave(file.path(save.dir, "./figs/beta/PCoA_fun_bray.pdf"),
#        pcoa_fun_plot, width = 89, height = 59, units = "mm")
pcoa_fun_plot
```
### Difference in functional variance among layers
```{r}
beta_fun_plot <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(fun_dist, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = Layers, y = distance)) + 
  geom_boxplot(width = 0.5, aes(fill = Layers)) +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_compare_means(comparisons = my_comparisons,  paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  #scale_fill_manual(values= cols)+
  labs(x = 'Layers', y = 'Functional variance', fill='Layers') +
  theme_bw() +
  theme(axis.title = element_text(size = 12, colour = "black"),
        axis.text = element_text(size = 10, colour = "black"),
        legend.title = element_text(size = 12),
        legend.text = element_text(size = 10),
        panel.grid = element_blank())
# ggsave(file.path(save.dir, "./figs/beta/fun_variance.pdf"),
#        beta_fun_plot, width = 89, height = 89, units = "mm")
beta_fun_plot
```
## Couple or decouple effect between taxonomic, phylogenetic and functional profile

```{r}
pcoa1_tax <- data.frame(Layers = metadata$Layer, scores(ord.tax)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
pcoa1_phy <- data.frame(Layers = metadata$Layer, ord.phy$vectors[, 1:2]) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
pcoa1_fun <- data.frame(Layers = metadata$Layer, scores(ord.fun)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
pcoa1_dat <- data.frame(pcoa1_tax[, 1:2], pcoa1_phy[,2], pcoa1_fun[,2])
colnames(pcoa1_dat) <- c("Layers", "Taxnomic_PCoA1", "Phylogenetic_PCoA1", "Functional_PCoA1")
```
Regression plot with PcoA axis using ggplot
```{r}
library(ggpmisc)
p_linear <- ggplot(pcoa1_dat, aes(x = Taxnomic_PCoA1, y = Functional_PCoA1, fill = Layers)) +
  geom_point(size=3.5, alpha=0.8, aes(colour = Layers)) +
  # geom_smooth(method="lm", size=1, se=T, colour='black') +
  stat_poly_line(colour='black') +
  stat_poly_eq(aes(label = paste(after_stat(eq.label),
                                 after_stat(rr.label), sep = "*\", \"*"), colour = Layers)) +
  scale_colour_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  ylab("Taxnomic PCoA1")+xlab("Functional PCoA1") +
  theme_bw() +
  theme(panel.grid=element_blank(), 
        axis.title = element_text(color = 'black', size = 6),
        axis.ticks.length = unit(0.2, "lines"), axis.ticks = element_line(color = 'black'),
        axis.line = element_blank(), 
        axis.text = element_text(colour = 'black',size = 5),
        strip.text = element_text(size = 6))
p_linear
```
Regression plot with dbRDA axis using ggplot
```{r}
dist_tax <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(tax_dist, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
dist_phy <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(beta.mntd.weighted, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
dist_fun <- sapply(unique(metadata$Layer), function(x) usedist::dist_subset(fun_dist, grep(x, metadata$sample_id, value = T))) %>%
  data.frame() %>% gather("Layers", "distance") %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
dist_dat <- data.frame(dist_tax[, 1:2], dist_phy[,2], dist_fun[,2])
colnames(dist_dat) <- c("Layers", "Taxnomic_distance", "Phylogenetic_distance", "Functional_distance")
p_linear <- ggplot(dist_dat, aes(x = Phylogenetic_distance, y = Functional_distance, fill = Layers)) +
  geom_point(size=3.5, alpha=0.8, aes(colour = Layers)) +
  geom_smooth(method="lm", size=1, se=T, colour='black') +
  scale_colour_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  stat_poly_line(colour='black') +
  stat_poly_eq(aes(label = paste(after_stat(eq.label),
                                 after_stat(rr.label), sep = "*\", \"*"), colour = Layers)) +
  ylab("Taxnomic distance")+xlab("Phylogenetic distance") +
  theme_bw() +
  theme(panel.grid=element_blank(), 
        axis.title = element_text(color='black',size=14),
        axis.ticks.length = unit(0.2,"lines"), axis.ticks = element_line(color='black'),
        axis.line = element_blank(), 
        axis.text.y = element_text(colour='black',size=12),
        axis.text.x = element_text(colour='black', size = 12),
        strip.text = element_text(size = 14),
        legend.position='none')
p_linear
```
## Composition
### Determine the OTU numbers within each dominant phylum
```{r}
# data preparation
otu_num_in_phylum <- data.frame(table(tax_table(phylo)[,"Phylum"]))
otu_num_in_phylum <- otu_num_in_phylum %>% arrange(desc(Freq))
otu_num_in_phylum <- rbind(otu_num_in_phylum[1:11, ], data.frame(Var1 = c('Others'), Freq = sum(otu_num_in_phylum[-c(1:11), 2])))

otu_num_in_phylum <- data.frame(Phylum = otu_num_in_phylum$Var1, otu_num = otu_num_in_phylum$Freq ,
                                prop = otu_num_in_phylum$Freq/sum(otu_num_in_phylum$Freq)*100)
otu_count.data <- otu_num_in_phylum %>%
  mutate(lab.ypos = cumsum(prop) - 0.5*prop)
otu_count.data$Phylum <- factor(otu_count.data$Phylum, ordered = T, levels = otu_num_in_phylum$Phylum)
```

```{r}
# pie plot 
# Define the colors you want
mycols <- c("#89c5da", "#ffc15c", "#74d944", "#CE50CA", "#5e738f", "#C0717C", "#CBD5ec", "#5F7FC7", 
                     "#00718b", "#00b0f6", "#a3a500", "#f8766d", "#D7C1B1", "#689030", "#AD6F3B", "#CD9BCD", 
                     "#D14285", "#6DDE88", "#652926", "#7FDCC0", "#C84248", "#8569D5", "#5E738F", "#D1A33D", 
                     "#8A7C64", "#599861")

pie_for_otu_num_phylum <- ggplot(otu_count.data, aes(x="", y=prop, fill=reorder(Phylum, -lab.ypos))) +
  geom_bar(width = 1, stat = "identity") + 
  coord_polar("y", start=0) +
  geom_text(aes(x = 1.35, y = lab.ypos, 
                label = paste0(otu_num, ' (', round(prop, 1), '%', ')', sep= '')),
            color = "black", size=3) +
  scale_fill_manual('Phylum', values = mycols) +
  guides(fill = guide_legend(reverse=T)) +
  theme_void() +
  theme(legend.position = "left")
pie_for_otu_num_phylum
```

### Determine the Order compositions within top 10 phyla
```{r}
arrange.tab <- function(phylo, N, taxrank, vect) {
  subphylo <- tax_glom(phylo, taxrank)
  subphylo.rel <- microbiome::transform(subphylo, "compositional")
  ra.tab <- otu_table(subphylo.rel)
  MRA <- rowMeans(ra.tab)
  group <- tax_table(subphylo.rel)[,vect]
  mra.tab <- data.frame(group,MRA)
  colnames(mra.tab) <- c('level1', 'level2', 'MRA')
  #arrange the class table
  mra.tab_level1 = mra.tab %>% group_by(level1) %>% 
    summarise(sum_MRA = sum(MRA)) %>% 
    arrange(desc(sum_MRA))
  top_N_level1 = mra.tab_level1[1:N, ]$'level1'
  top_N_tab = mra.tab[mra.tab$'level1' %in% top_N_level1, ]
  mra.tab_level2 = top_N_tab %>% group_by(level2) %>% 
    summarise(sum_MRA = sum(MRA)) %>% 
    arrange(desc(sum_MRA))
  order_level2 = mra.tab_level2$'level2'
  top_N_tab$'level1' = factor(top_N_tab$'level1', ordered = T, levels = top_N_level1)
  top_N_tab$'level2' = factor(top_N_tab$'level2', ordered = T, levels = rev(order_level2))
  top_N_tab$MRA = top_N_tab$MRA*100
  return(top_N_tab)
}
top10phylum <- arrange.tab(phylo, 10, 'Order', c(2,4))
mra.tab_level1 = top10phylum %>% group_by(level1) %>% 
  summarise(sum_MRA = sum(MRA)) %>% 
  arrange(desc(sum_MRA))
mra.tab_level1
mra.tab_level2 = top10phylum %>% group_by(level2) %>% 
  summarise(sum_MRA = sum(MRA)) %>% 
  arrange(desc(sum_MRA))
mra.tab_level2 [1:20, ]
order_level2 = mra.tab_level2$'level2'
```

```{r, fig.width = 8, fig.height = 6}
#stack bar plot
taxa_barplot <- ggplot(top10phylum, aes(fill=level2, y=MRA, x=level1)) + 
  geom_bar(position="stack", stat="identity") +
  scale_fill_manual('Order', breaks = order_level2[1:15], 
                    values = rep(c(rev(mycols[1:12]), mycols[-c(1:12)]), 20)[1:nrow(top10phylum)]) + #only the top 10 phylum and top 10 order are showed
  labs(x = 'Phylum', y = 'Mean relative abundance (%)') +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 30)) +
  theme_classic()+
  theme(legend.position = c(0.6,0.6),
        panel.grid=element_blank(), 
        axis.title = element_text(color = 'black', size = 8),
        axis.ticks.length = unit(0.4, "lines"), axis.ticks = element_line(color = 'black'),
        axis.line = element_line(colour = "black"), 
        axis.text.y = element_text(colour = 'black', size = 6),
        axis.text.x = element_text(colour = 'black', size = 6, angle = 45, hjust = 1),
        legend.text = element_text(size = 6),
        legend.key = element_blank(),
        legend.background = element_rect(colour = "white"))
taxa_barplot
```

### Determine the compositions within top 10 phylums for each layer
```{r}
subphylo <- tax_glom(phylo, 'Phylum')
subphylo.rel  = transform_sample_counts(subphylo, function(x) x / sum(x) )
ntaxa(subphylo.rel)
ra.tab <- otu_table(subphylo.rel)
sum(ra.tab[, 1])
subtaxa_tab <- tax_table(subphylo.rel)[, 2]
data.frame(subtaxa_tab, ra.tab) %>% group_by(Phylum) %>% 
  summarise(across(everything(), sum)) %>% 
  mutate(MRA = rowMeans(select(., colnames(ra.tab)))) %>%
  arrange(desc(MRA)) %>% dplyr::top_n(10, MRA) %>%
  select(., c('Phylum', 'MRA'))
```

```{r, fig.width = 8, fig.height = 4.5}
## boxplot shows the community composition for each layer
box_plot <- data.frame(subtaxa_tab, ra.tab) %>% group_by(Phylum) %>% 
  summarise(across(everything(), sum)) %>% 
  mutate(MRA = rowMeans(select(., colnames(ra.tab)))) %>%
  arrange(desc(MRA)) %>% dplyr::top_n(10, MRA) %>%
  select(., -c('MRA')) %>% 
  mutate(Phylum = factor(Phylum, levels = Phylum)) %>%
  tidyr::pivot_longer(cols = -c(Phylum), names_to = "sample_id", values_to = 'rel_abun') %>%
  mutate(layer = sapply(stringr::str_split(sample_id, "_",  n = 2), `[`, 1)) %>%
  mutate(layer = factor(layer, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(Phylum, rel_abun*100)) + 
  geom_boxplot(width = 0.5, aes(fill = layer)) +
  stat_compare_means(aes(group = layer),  paired = TRUE, 
                     p.adjust.method = "BH", label = "p.signif") +
  scale_fill_manual(values = c("#f8766d", "#a3a500", "#00b0f6")) +
  labs(x = 'Phylum', y = 'Relative abundance (%)', fill='Layers') +
  theme_bw() +
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  theme(axis.title = element_text(size = 8, colour = "black"),
        axis.text.x = element_text(size = 6, colour = "black", 
                                   angle = 45, vjust = 1, hjust = 1),
        axis.text.y = element_text(size = 6, colour = "black"),
        legend.title = element_text(size = 8),
        legend.text = element_text(size = 6),
        panel.grid = element_blank())
box_plot
```
### Barplot shows the community composition
```{r, fig.width = 8, fig.height = 4.5}
bar_plot <- data.frame(subtaxa_tab, ra.tab) %>% 
  mutate(MRA = rowMeans(select(., colnames(ra.tab)))) %>%
  arrange(desc(MRA)) %>% dplyr::top_n(10, MRA) %>%
  select(., -c('MRA')) %>% 
  bind_rows(summarise_all(., ~if(is.numeric(.)) 1-sum(.) else "Other")) %>%
  mutate(Phylum = factor(Phylum, levels = Phylum)) %>%
  pivot_longer(cols = -c(Phylum), names_to = "sample_id", values_to = 'rel_abun') %>%
  mutate(layer = sapply(stringr::str_split(sample_id, "_",  n = 2), `[`, 1)) %>%
  mutate(layer = factor(layer, levels = c('SUR', 'SUB', 'PL'))) %>%
  ggplot(aes(x = sample_id, y = 100*rel_abun, fill = Phylum))+
  geom_bar(stat = "identity") +
  scale_fill_brewer(palette = "Paired")+
  scale_y_continuous(expand = c(0,0))+
  labs(x = 'Sample', y = 'Mean relative abundance (%)') +
  theme_linedraw() + 
  guides(fill = guide_legend(keywidth = 0.5, keyheight = 0.5)) +
  facet_grid(~layer, scales = "free_x", space = "free_x") + 
  theme(axis.title.x = element_blank(), 
        axis.title.y = element_text(size = 8, colour = "black"),
        panel.grid.minor = element_blank(), 
        panel.grid.major = element_blank(), 
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1, size = 6), 
        axis.text.y = element_text(size = 6), 
        panel.spacing = unit(0, "lines"),
        strip.text = element_blank(),
        strip.background = element_blank(),
        legend.title = element_text(size = 8), 
        legend.text = element_text(size = 6))
# ggsave(file.path(save.dir, "./figs/composition/tax_composition_10.pdf"), box_plot, width = 189, height = 95, units = "mm")  
# ggsave(file.path(save.dir, "./figs/composition/tax_composition_bar.pdf"), bar_plot, width = 120, height = 65, units = "mm") 
bar_plot
```
### Unique otus profile among three layers
```{r, fig.width = 6, fig.height = 4.5}
# Load the library
library(limma)
# Generate example data
sur_venn <- otu %>% data.frame() %>%
  mutate(rowsum = rowSums(select(., grep('SUR', metadata$sample_id, value = T)))) %>%
  filter(rowsum > 0) %>%
  rownames()
sub_venn <- otu %>% data.frame() %>%
  mutate(rowsum = rowSums(select(., grep('SUB', metadata$sample_id, value = T)))) %>%
  filter(rowsum > 0) %>%
  rownames()
pl_venn <- otu %>% data.frame() %>%
  mutate(rowsum = rowSums(select(., grep('PL', metadata$sample_id, value = T)))) %>%
  filter(rowsum > 0) %>%
  rownames()

# What are the possible letters in the universe?
universe <- sort(unique(c(sur_venn, sub_venn, pl_venn)))

# Generate a matrix, with the sets in columns and possible letters on rows
Counts <- matrix(0, nrow=length(universe), ncol=3)
# Populate the said matrix
for (i in 1:length(universe)) {
  Counts[i,1] <- universe[i] %in% sur_venn
  Counts[i,2] <- universe[i] %in% sub_venn
  Counts[i,3] <- universe[i] %in% pl_venn
}

# Specify the colors for the sets
# Prepare a palette of 3 colors:
gg_color_hue <- function(n) {
  hues = seq(15, 375, length = n + 1)
  hcl(h = hues, l = 65, c = 100)[1:n]
}
n = 3
cols = gg_color_hue(n)
# Name the columns with the sample names
colnames(Counts) <- c("SUR","SUB","PL")
vennDiagram(vennCounts(Counts), circle.col = cols)
```
Venn plot using VennDiagram
```{r}
# library
library(VennDiagram)

#Make the plot
venn_plot <- venn.diagram(
  x = list(
    otu %>% data.frame() %>%
      mutate(rowsum = rowSums(select(., grep('SUR', metadata$sample_id, value = T)))) %>%
      filter(rowsum > 0) %>%
      rownames(), 
    otu %>% data.frame() %>%
      mutate(rowsum = rowSums(select(., grep('SUB', metadata$sample_id, value = T)))) %>%
      filter(rowsum > 0) %>%
      rownames(), 
    otu %>% data.frame() %>%
      mutate(rowsum = rowSums(select(., grep('PL', metadata$sample_id, value = T)))) %>%
      filter(rowsum > 0) %>%
      rownames()
  ),
  category.names = c("SUR" , "SUB" , "PL"),
  filename = NULL,
  # output = F,
  # imagetype="png" ,
  # height = 480 ,
  # width = 480 ,
  # resolution = 300,
  # compression = "lzw",
  # Circles
  lwd = 2,
  lty = 'blank',
  fill = cols,
  # Numbers
  cex = 1,
  fontfamily = "sans",
  # Set names
  cat.cex = 1,
  cat.default.pos = "outer",
  cat.pos = c(-27, 27, 135),
  cat.dist = c(0.055, 0.055, 0.085),
  cat.fontfamily = "sans",
  cat.col = 'black',
  rotation = 1
)
# grid.newpage()
# pdf(file= file.path(save.dir, "OTU_Group_venn.pdf"), height = 3, width = 3)
# grid.draw(venn_plot)
# dev.off()
grid::grid.draw(venn_plot)
```
