---
title: "Environmetal effects on diversity and composition"
author: "Luyao Kang"
date: "`r Sys.Date()`"
output: html_notebook
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
```{r setup, include = F}
knitr::opts_chunk$set(
  collapse = T, echo = T, comment = "#>", message = F, warning = F,
	fig.align = "center", fig.width= 4.5, fig.height = 3, dpi = 150)
```
## Data input 
Set work directory
```{r}
setwd('e:/permafrost/')
wd_16s <- file.path(getwd(),"data/16S")
# if (!dir.exists(wd_16s)) {
#   dir.create(wd_16s)
# }
wd_fun <- file.path(getwd(),"data/metagenome")
save.dir <- file.path(getwd(),"result")
```
Read data
```{r}
source("script/read_data.R")
```
## Test the environmetanl effects on the microbial diversity
Loading packages
```{r}
library(ggplot2)
library(tidyr)
library(dplyr)
library(tibble)
```
Prepare the table for correlation
```{r}
alpha_div <- estimate_richness(phylo_rare, measures = c("Observed", "Chao1", 'Shannon', 'Simpson'))
library(picante)
pd <- pd(t(otu), tree, include.root = F) # estimate the phylogenetic diversity
alpha_div <- cbind(Layers = metadata$Layer, alpha_div, Faith = pd$PD,
                   Evenness = alpha_div$Shannon/log(alpha_div$Observed)) %>%
  mutate(Layers = factor(Layers, levels = c('SUR', 'SUB', 'PL')))
cor_div_env_tab <- data.frame(metadata[-c(1, 2, 4:6)], Evenness = alpha_div$Evenness)
# You can use sel_env to specify the variables you want to use and sel_env_label to specify the labes for the pannel
sel_env <- c("MAT", "AI", "Moisture",	"pH",	"Clay",	"Silt",	"Sand",	"TC", 
           "SOC", "DOC", "TN",	"NH4_N",	"NO3_N",	"DON",	
           "LCP1",	"LCP2",	"RCP")
```
Correlation analysis
```{r}
groups <- cor_div_env_tab$Layer
x <- data.frame(cor_div_env_tab[, sel_env])
y <- data.frame(Evenness = cor_div_env_tab$Evenness)
# You can use kendall, spearman, or pearson below:
method<-"pearson"
# Now calculate the correlation between individual Taxa and the environmental data
df <- NULL
for (i in colnames(x)) {
  for (j in colnames(y)) {
    for (k in unique(groups)) {
      a <- x[groups == k, i, drop = F]
      b <- y[groups == k, j, drop = F]
      tmp <- c(i, j, cor(a[complete.cases(b), ], b[complete.cases(b), ], use = "everything", method = method), 
               cor.test(a[complete.cases(b), ], b[complete.cases(b), ], method = method)$p.value, k)
      if (is.null(df)) {
        df <- tmp
      } else {
        df <- rbind(df, tmp)
      }
    }
  }
}

df <- data.frame(row.names = NULL, df)
colnames(df) <- c("Env", "div", "r", "Pvalue", "Layer")
df$Pvalue <- as.numeric(as.character(df$Pvalue))
df$AdjPvalue <- rep(0, dim(df)[1])
df$r <- as.numeric(as.character(df$r))

# Now we generate the labels for signifant values
df$Significance <- cut(df$Pvalue, breaks = c(-Inf, 0.001, 0.01, 0.05, Inf), label = c("***", "**", "*", ""))

# We ignore NAs
df <- df[complete.cases(df), ]
df <- df[, c("Env", "Layer", "r", "Significance")]
# We want to reorganize the Env data based on they appear
df$Env <- factor(df$Env, levels = sel_env)
df$Layer <- factor(df$Layer, levels = rev(c("SUR", "SUB", "PL")))
```
Heatmap for alpha diversity and environmental factors
```{r, fig.align='center', fig.width=6.5, fig.height=6.5}
row_num = length(levels(df$Layer))
p <- ggplot(aes(x = Env, y = as.numeric(Layer), fill = r), data = df) + 
  xlim(c("", sel_env)) +
  ylim(c(-row_num/1.5, row_num + 1)) +
  geom_tile() + ylab("") + 
  scale_fill_gradient2(low = "#2C7BB6", mid = "white", high = "#D7191C", limits = c(-1, 1)) + 
  annotate(x = "", y = 1:row_num, label = levels(df$Layer), size = 2.5, geom = "text") +
  geom_text(aes(label = Significance), color = "black", size = 4)

p <- p + coord_polar(start = -0.15, theta = "x") + 
  theme_bw() + 
  theme(axis.text = element_text(color = "black"),
        legend.position = c(0.5, 0.5), 
        legend.key.size = unit(0.5, "cm"))
# ggsave(file.path(save.dir, './figs/env_effect/env_alpha_plot.pdf'), p, width = 9, height = 6, units = "in")
p
```
## Test the environmental effects on the community structure
Partial mantel test function for the relationship between the envs and community
```{r}
partial.mantel.fun <- function(phylo) {
  env.table <- data.frame(sample_data(phylo))
  #env.table <- env.table[complete.cases(env.table), ]
  otu_table <- as.matrix(t(otu_table(phylo)))
  otu_table_hel <- decostand(otu_table, 'hellinger')
  otu_table_hel_dist <- vegdist(otu_table_hel, 'bray',upper=F)
  df <- NULL
  vars <- c("MAT", "AI", "Moisture",	"pH",	"Clay",	"Silt",	"Sand",	"TC", 
            "SOC", "DOC", "TN",	"NH4_N",	"NO3_N",	"DON",	
            "LCP1",	"LCP2",	"RCP")
  for (x in vars) {
    y.dist <- vegdist(scale(env.table[,x]), 'euclidean', na.rm = T)
    z.dist <- vegdist(scale(env.table[ , setdiff(vars, x)]), 'euclidean', na.rm = T)
    mode <- mantel.partial(otu_table_hel_dist, y.dist, z.dist, 
                           method = "pearson", permutations = 999, na.rm = T)
    r <- mode$statistic
    p <- mode$signif
    tmp <- data.frame(env = x, r = r, p.value = p)
    if(is.null(df))
      df <- tmp
    else
      df <- rbind(df ,tmp)
  }
  return(df)
}

phylo_sur <- subset_samples(phylo, Layer == 'SUR')
phylo_sur <- prune_taxa(taxa_sums(phylo_sur)>=1, phylo_sur)
phylo_sub <- subset_samples(phylo, Layer == 'SUB')
phylo_sub <- prune_taxa(taxa_sums(phylo_sub)>=1, phylo_sub)
phylo_pl <- subset_samples(phylo, Layer == 'PL')
phylo_pl <- prune_taxa(taxa_sums(phylo_pl)>=1, phylo_pl)

set.seed(123)
par.mant.sur <- partial.mantel.fun(phylo_sur)
set.seed(123)
par.mant.sub <- partial.mantel.fun(phylo_sub)
set.seed(123)
par.mant.pl <- partial.mantel.fun(phylo_pl)
```
PLOT using ggcor package
```{r, fig.align='center', fig.width=8, fig.height=5}
## devtools::install_github('hannet91/ggcor')
library(ggcor)
par.man.tibble <- tibble(spec = c(rep('SUR', nrow(par.mant.sur)), rep('SUB', nrow(par.mant.sub)), rep('PL', nrow(par.mant.pl))), 
                         rbind(par.mant.sur, par.mant.sub, par.mant.pl))

vars <- c("MAT", "AI", "Moisture",	"pH",	"Clay",	"Silt",	"Sand",	"TC", 
          "SOC", "DOC", "TN",	"NH4_N",	"NO3_N",	"DON",	
          "LCP1",	"LCP2",	"RCP")
env.table <- sample_data(phylo)[ , vars]
mantel02 <- par.man.tibble %>% 
  mutate(r = cut(r, breaks = c(-Inf, 0.2, 0.5, Inf), 
                 labels = c("<0.20", "0.20-0.5", ">0.50"),
                 right = FALSE),
         p.value = cut(p.value, breaks = c(-Inf, 0.001, 0.01, 0.05, Inf),
                       labels = c("<0.001", "0.001-0.01", "0.01-0.05", ">0.05"),
                       right = T))
p <- quickcor(env.table, type = "lower") + geom_square() + 
  add_link(mantel02, mapping = aes(colour = p.value, size = r),
           diag.label = TRUE) +
  scale_color_manual(values = c('#d95f02', '#1b9e77', '#3C5488FF', 'grey')) +
  scale_size_manual(values = c(0.5, 2, 3)) +
  geom_diag_label() + remove_axis("x")
# ggsave(file.path(save.dir, './figs/env_effect/partail_matel_plot.pdf'), p, width = 9, height = 6, units = "in")
p
```
